---
title: "FD project -  Spatial analyses"
author: "Brunno F Oliveira^1^; Brett R Scheffers^1^;"
date: "October 20, 2015\n**1** Department of Wildlife Ecology and Conservation, University of Florida/IFAS, Gainesville, FL 32611, USA. **Corresponding author:** (brunno.oliveira@me.com) "
output: pdf_document
theme: united
---

*** 

\newpage

# Packages versions:
```{r info, message=FALSE,echo=F}
info <- sessionInfo()
```

We used `r info[1]$R$ver` and the following packages:

```{r packages, message=FALSE, echo=T}
rm(list=ls())
gc()

list.of.packages <- c("raster","spdep","ncf","doParallel","foreach","piecewiseSEM","nlme","plotrix","QuantPsyc","pgirmess","MuMIn","ggplot2","SpatialPack","mctest","GGally","maps","data.table","DiagrammeR","V8","DiagrammeRsvg","htmltools","rsvg","Rgraphviz","knitr","PerformanceAnalytics","spatialreg")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

sapply(list.of.packages, require, character.only = TRUE)

```

# Set working directory
```{r setup, cache = FALSE, include=FALSE}
require("knitr")

savedatafolder <- "C:/Users/boliveir/Dropbox (UFL)/FD_causes/" # address to folder to save results
externHD <- 'G:/GIS/' # address to external HD
setwd("C:/Users/boliveir/Dropbox (UFL)/FD_causes")
opts_knit$set(root.dir = "C:/Users/boliveir/Dropbox (UFL)/FD_causes")

getwd()

#### function to get equal binds from a vector
func_splint <- function(x,interval=4) {
  require(ggplot2)
  is.odd <- function(x) x %% 2 != 0
  
  a <- levels(cut_interval(x,interval-1))
  b <- unlist(strsplit(a,','))
  c <- gsub('[','',b,fixed="TRUE")
  d <- gsub(']','',c,fixed="TRUE")
  e <- gsub('(','',d,fixed="TRUE")
  f <- gsub(')','',e,fixed="TRUE")
  return(as.numeric(c(unique(f))))
}
```

## Land data
```{r , echo=F}
land <- rgdal::readOGR("G:/GIS/Shp files/ne_50m_land/ne_50m_land_no_artic.shp")
crs(land) <-"+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 "
land <- spTransform(land, CRS("+proj=moll +datum=WGS84"))

```

# Load data
```{r, echo = FALSE}
taxa <- c("amphibians","birds","mammals","reptiles")

metadata <- list()
for(i in 1:length(taxa)){
  metadata[[i]] <- read.csv(paste(savedatafolder,"results/",taxa[i],"_metadata.csv", sep = ""))[,-1]
}
names(metadata) <- taxa

XY <- list()
for(i in 1:length(taxa)){
  XY[[i]] <- metadata[[i]][,1:2]
  XY[[i]] <- XY[[i]]/1000 # convert coordinates from meters to km
}
names(XY) <- taxa


```

# Check pairwise correlation among FD metrics
```{r}
FD_metrics <- c("FDis","ses.FD.realm","ses.FD.biome","ses.FD.biorealm")

# For amphibians
pairs(metadata$amphibians[,FD_metrics])
# we identify outliers in ses.FD.realm
# boxplot(metadata$amphibians$ses.FD.realm)
# we therefore decided for removing these outliers
# metadata$amphibians$ses.FD.realm[metadata$amphibians$ses.FD.realm < (-70)] <- NA
# metadata$amphibians$ses.FD.realm[metadata$amphibians$ses.FD.realm < (-60)] <- NA

# For birds
pairs(metadata$birds[,FD_metrics])
# we identify outliers in ses.FD.realm
# boxplot(metadata$birds$ses.FD.realm)
# we therefore decided for removing these outliers
# metadata$birds$ses.FD.realm[metadata$birds$ses.FD.realm < (-220)] <- NA

# For mammals
pairs(metadata$mammals[,FD_metrics])
# we identify outliers in ses.FD.realm
# boxplot(metadata$mammals$ses.FD.realm)
# we therefore decided for removing these outliers
# metadata$mammals$ses.FD.realm[metadata$mammals$ses.FD.realm < (-100)] <- NA

# For reptiles
pairs(metadata$reptiles[,FD_metrics])
# we identify outliers in ses.FD.realm
# boxplot(metadata$reptiles$ses.FD.realm)
# we therefore decided for removing these outliers
# metadata$reptiles$ses.FD.realm[metadata$reptiles$ses.FD.realm < (-50)] <- NA
```

# Correlations
```{r warning= F}

for(i in 1:length(taxa)){
  png(paste(savedatafolder,"results/",taxa[i],"_correlations.png",sep=""),
      units = "in", width = 7, height = 7, res = 300)
  chart.Correlation(data.frame(
    richness = metadata[[i]]$SPD,
    Div = metadata[[i]]$DIV_eq,
    Div_ses = metadata[[i]]$ses_DIV_eq_realm,
    MPD = metadata[[i]]$MPD_,
    MPD_ses = metadata[[i]]$ses_MPD_realm,
    FDis_ses = metadata[[i]]$ses.FD.realm,
    velocity = metadata[[i]]$velocity, 
    Temp_stab = metadata[[i]]$temp.stab, 
    Temp_dist = metadata[[i]]$tempdist, 
    Temp_nov = metadata[[i]]$tempnov, 
    Clim_stab = metadata[[i]]$climate.stab, 
    Clim_dist = metadata[[i]]$climdist, 
    Clim_nov = metadata[[i]]$climnov), 
    histogram=TRUE, pch=19)
  dev.off()
}

# remove outilers for FDses in reptiles
# hist(metadata$reptiles$ses.FD.realm)
# metadata$reptiles$ses.FD.realm[which(metadata$reptiles$ses.FD.realm < -50)] <- NA

# remove outilers for DIVses in reptiles
# hist(metadata$reptiles$ses_DIV_eq_realm)
# metadata$reptiles$ses_DIV_eq_realm[which(metadata$reptiles$ses_DIV_eq_realm > 7)] <- NA

# remove outilers for MPDses in reptiles
# hist(metadata$reptiles$ses_MPD_realm)
# metadata$reptiles$ses_MPD_realm[which(metadata$reptiles$ses_MPD_realm < -14)] <- NA

# remove outilers for FDses in birds
# hist(metadata$birds$ses.FD.realm)
# metadata$birds$ses.FD.realm[which(metadata$birds$ses.FD.realm < -110)] <- NA

```

# Create datasets
```{r}

# Global datasets

dataset.group <- list()

for(i in 1:length(taxa)){
  
  metadata[[i]]$ses.FD.realm[
    which(is.infinite(metadata[[i]]$ses.FD.realm))] <- NA
  
  dataset.tmp <- data.frame(na.omit(
    data.frame(XY[[i]],
               Realm = as.character(metadata[[i]]$Realm),
               scale(data.frame(
                 SPD = log(metadata[[i]]$SPD+1),
                 FD = metadata[[i]]$ses.FD.realm,
                 Div = metadata[[i]]$ses_DIV_eq_realm, 
                 Age = metadata[[i]]$ses_MPD_realm, 
                 Stab = metadata[[i]]$temp.stab, 
                 Prod = metadata[[i]]$NPP, 
                 TRI = metadata[[i]]$tri,
                 Temperature = metadata[[i]]$mean.temp)))))
  
  dataset.group[[i]] <- dataset.tmp
}
names(dataset.group) <- taxa

### Regional datasets
Realms <- as.character(na.omit(unique(metadata[[2]]$Realm)))
Realms <- Realms[-which(Realms=="Oceania")]
NRealm <- length(Realms)

reg_dataset <- list()

for (i in 1:length(taxa)){
  reg_dataset.tmp <- list()
  for (j in 1:NRealm){
    
    reg_dataset.tmp_ <- dataset.group[[i]]
    
    reg_dataset.tmp[[j]] <- reg_dataset.tmp_[which(as.character(reg_dataset.tmp_$Realm) == Realms[j]),]
    
  }
  names(reg_dataset.tmp) <- Realms
  reg_dataset[[i]] <- reg_dataset.tmp
  
}
names(reg_dataset) <- taxa

# Stab = max(metadata[[i]]$temp.stab) - metadata[[i]]$temp.stab , # invert temp.stab  (higher values are more stable, not instable)
```


# Create Raster layer stack
```{r}
s <- list()
for(i in 1:length(taxa)){
  s.temp <- stack()
  temp.meta <- metadata[[i]]
  for(n in 3:ncol(temp.meta)){ # dont use X and Y
    d <- rasterFromXYZ(data.frame(x=temp.meta[,1],
                                  y=temp.meta[,2],
                                  z=temp.meta[,n]),
                       crs=CRS("+proj=cea +datum=WGS84"))
    
    d <- projectRaster(d, crs = crs(land)) # Project to Mollweide
    
    s.temp <- addLayer(s.temp,d)
  }
  names(s.temp) <- names(temp.meta)[3:ncol(temp.meta)]
  s[[i]] <- s.temp
}

names(s) <- taxa

```

# Maps
## Prepare
```{r getJenksBreaks, warning= F}
maps_vars <- c("SPD", "ses_DIV_eq_realm", "ses_MPD_realm", "ses.FD.realm")

class_int <- list()
for(i in 1:length(taxa)){
  class_int.tmp <- list()
  for(j in 1:length(maps_vars)){
    # Get class intervals
    tmp <- na.omit(metadata[[i]][,maps_vars[j]])
    tmp[is.infinite(tmp)] <- NA
    class_int.tmp[[j]] <- 
      BAMMtools::getJenksBreaks(tmp, 20)
    # avoid identical breaks
    class_int.tmp[[j]] <- unique(class_int.tmp[[j]])
  }
  names(class_int.tmp) <- maps_vars
  class_int[[i]] <- class_int.tmp
}
names(class_int) <- taxa

```


## Plot maps
```{r maps, warning= F}

maps_vars_names <- c("Species richness", 
                     "Diversification rates",
                     "Assemblage age (SES)",
                     "Functional diversity (SES)")

png(paste(savedatafolder,"results/maps.png",sep=""), 
    units = "in", width = 11, height = 5, res = 300)
layout(matrix(c(1:(4*4)), 4, 4, byrow = T))

par(mar=c(0,0,2,0))

for(i in 1:length(maps_vars)){
  for(j in 1:length(taxa)){
    if(i == 1){
      maintitle <- paste(toupper(substr(taxa[j], 1, 1)), 
                         substr(taxa[j], 2, nchar(taxa[j])),"\n", 
                         maps_vars_names[i], sep="")
      labs <- round(func_splint(class_int[[j]][[maps_vars[i]]],5),0)
    }
    else{
      maintitle <- paste("\n",maps_vars_names[i], sep="")
      labs <- round(func_splint(class_int[[j]][[maps_vars[i]]],5),0)
    }
    if(i == 2 | i == 3){ # for SES zero should be in the middle
      maintitle <- paste("\n",maps_vars_names[i], sep="")
      labs <- round(func_splint(class_int[[j]][[maps_vars[i]]],5),2)
    }
    
    
    plot(land, main=maintitle, cex=.5, border="white")
    plot(s[[j]][[maps_vars[i]]],add=T,
         axes=F,box=F,legend=FALSE,
         axis.args=list(at=labs,
                        labels=labs), 
         breaks = class_int[[j]][[maps_vars[i]]],
         col=colorRampPalette(c('blue','lightblue','yellow','red'))(20))
    map(land,add=T,cex=.5)
    plot(s[[j]][[maps_vars[i]]],add=T,main='',axes=F,box=F,
         legend.only=TRUE,smallplot=c(0.05,.07, 0.05,.5),
         axis.args=list(at=labs,
                        labels=labs), 
         breaks = class_int[[j]][[maps_vars[i]]],
         col=colorRampPalette(c('blue','lightblue','yellow','red'))(20))
  }
}

dev.off()

```

## Plot FD maps
```{r maps, warning= F}

png(paste(savedatafolder,"results/FD_maps.png",sep=""), 
    units = "in", width = 10, height = 5, res = 300)
layout(matrix(c(1:4), 2, 2, byrow = T))

i = 4

par(mar=c(0,0,1,0))

for(j in 1:length(taxa)){
  
  maintitle <- paste(toupper(substr(taxa[j], 1, 1)), 
                     substr(taxa[j], 2, nchar(taxa[j])), sep="")
  labs <- round(func_splint(class_int[[j]][[maps_vars[i]]],5),0)
  
  plot(land, main=maintitle, cex=.5, border="white")
  plot(s[[j]][[maps_vars[i]]],add=T,
       axes=F,box=F,legend=FALSE,
       axis.args=list(at=labs,
                      labels=labs), 
       breaks = class_int[[j]][[maps_vars[i]]],
       col=colorRampPalette(c('blue','lightblue','yellow','red'))(20))
  map(land,add=T,cex=.5)
  plot(s[[j]][[maps_vars[i]]],add=T,main='',axes=F,box=F,
       legend.only=TRUE,smallplot=c(0.05,.07, 0.05,.5),
       axis.args=list(at=labs,
                      labels=labs), 
       breaks = class_int[[j]][[maps_vars[i]]],
       col=colorRampPalette(c('blue','lightblue','yellow','red'))(20))
}

dev.off()

```




# Test correlation among predictors
```{r warning= F}
independents <- 
  data.frame(scale(data.frame(velocity = metadata[[i]]$velocity, 
                              Temp_stab = metadata[[i]]$temp.stab, 
                              Temp_dist = metadata[[i]]$tempdist, 
                              Temp_nov = metadata[[i]]$tempnov, 
                              Clim_stab = metadata[[i]]$climate.stab, 
                              Clim_dist = metadata[[i]]$climdist, 
                              Clim_nov = metadata[[i]]$climnov,
                              NPP = metadata[[i]]$NPP,
                              TRI = metadata[[i]]$tri,
                              Mean_Temp = metadata[[i]]$mean.temp)))

#View(cor(na.omit(independents)))
#usdm::vif(data.frame(scale(independents)))
usdm::vifstep(scale(independents), th=3)
independents_vif_vars <- usdm::vifstep(scale(independents), th=3)
independents_vif <- independents[,-which(
  names(independents) %in% independents_vif_vars@excluded)]
usdm::vif(data.frame(scale(independents_vif)))
#View(cor(na.omit(independents_vif)))

# We choose the following predictors:
# ses.MPD
# Temperature distance
# mean temperature
# TRI
# We removed temperature novelty, temperature stability and velocity because they all showed a smaller correlation with Div and FD.ses than temperature distance. The exception was reptiles, that showed strongest correlation between temperature stability and Div and FD.ses.

```



# OLS model
## Global models
```{r warning= F}

options(na.action("na.omit"))

lm.models <- list()

for(i in 1:length(taxa)){
  
  model.div <- lm(Div ~ 
                    Age+
                    Stab,
                  data = dataset.group[[i]])
  
  model.div2 <- lm(Div ~ 
                    Age+
                    Stab,
                  data = dataset.group[[i]])
  
  model.age <- lm(Age ~ 
                    Stab,
                  data = dataset.group[[i]])
  
  model.FD <- lm(FD ~ 
                   Div+
                   Age+
                   Stab+
                   Prod,
                 data = dataset.group[[i]])
  
  lm.models[[i]] <- list(Div = model.div,
                         Div2 = model.div2,
                         Age = model.age,
                         FD = model.FD)
}


```

## Run Regional models
```{r}

res.lm.rich.realm <- list()
res.lm.div.realm <- list()
res.lm.FD.realm <- list()

## Models by realm
model.summary.realm <- list(NA)

for(i in 1:length(taxa)){ 
  model.summary.realm_ <- list()
  
  res.lm.rich.realm_ <- list()
  res.lm.div.realm_ <- list()
  res.lm.FD.realm_ <- list()
  
  for(j in 1:NRealm){ 
    cat("Taxa", taxa[i], '-', i, 'from', length(taxa), 
        "\n Realm", Realms[i], '-', i, 'from', NRealm,"\r")
    
    datatmp <- reg_dataset[[i]][[j]]
    
    model.rich <- lm(SPD ~ 
                       FD+
                       Div+
                       Age+
                       Stab+
                       Prod+
                       TRI+
                       Temperature,
                     data = datatmp, 
                     na.action=na.fail)
    
    model.div <- lm(Div ~ 
                      Age+
                      Stab+
                      #Prod+
                      TRI+
                      Temperature,
                    data = datatmp, 
                    na.action=na.fail)
    
    model.FD <- lm(FD ~ 
                     Div+
                     Age+
                     Stab+
                     Prod,
                   data = datatmp, 
                   na.action=na.fail)
    
    res.lm.rich.realm_[[j]] <- residuals(model.rich)
    res.lm.div.realm_[[j]] <- residuals(model.div)
    res.lm.FD.realm_[[j]] <- residuals(model.FD)
    
    model.rich.2 <- summary(model.rich)
    model.rich.summary <- model.rich.2$coefficients[-1,]
    
    model.div.2 <- summary(model.div)
    model.div.summary <- model.div.2$coefficients[-1,]
    
    model.FD.2 <- summary(model.FD)
    model.FD.summary <- model.FD.2$coefficients[-1,]
    
    model.summary_ <- data.frame(
      rbind(data.frame(model = "Rich",
                       data.frame(
                         var = rownames(model.rich.summary),
                         model.rich.summary),
                       confint(model.rich)[-1,]),
            data.frame(model = "Div",
                       data.frame(
                         var = rownames(model.div.summary),
                         model.div.summary),
                       confint(model.div)[-1,]),
            data.frame(model = "FD",
                       data.frame(
                         var = rownames(model.FD.summary),
                         model.FD.summary),
                       confint(model.FD)[-1,])))
    
    names(model.summary_)[7] <- "Lower_IC"
    names(model.summary_)[8] <- "Higher_IC"
    
    model.summary_$sig <- ifelse(model.summary_$Pr...t..<0.05,1,0)
    
    model.summary.realm_[[j]] <- model.summary_
    
  }
  names(model.summary.realm_) <- Realms
  
  model.summary.2.realm <- rbindlist(model.summary.realm_,idcol = T)
  model.summary.realm[[i]] <- model.summary.2.realm
  
  names(res.lm.rich.realm_) <- Realms
  names(res.lm.div.realm_) <- Realms
  names(res.lm.FD.realm_) <- Realms
  
  res.lm.rich.realm[[i]] <- res.lm.rich.realm_
  res.lm.div.realm[[i]] <- res.lm.div.realm_
  res.lm.FD.realm[[i]] <- res.lm.FD.realm_
}

names(model.summary.realm) <- taxa

model.summary.realm <- rbindlist(model.summary.realm,idcol = T)
names(model.summary.realm)[1:2] <- c("taxa", "realm")

names(res.lm.rich.realm) <- taxa
names(res.lm.div.realm) <- taxa
names(res.lm.FD.realm) <- taxa

# add global model
model.summary.2$REALM <- "Global"
model.summary.2 <- rbind(model.summary.2,model.summary.realm)

```


## Plot model coeffs OLS models
```{r warning= F}
col_realms <- c("#800000", "#808000", "#000075",
                "#e6194B", "#f58231", "#3cb44b", "#42d4f4",
                "#911eb4", "#fabebe", "#ffd8b1", "#e6beff")
names(col_realms) <- c(Realms)

pdf(paste(savedatafolder,"results/OLS_models.pdf",sep=""), width = 10, height = 5)
#png(paste(savedatafolder,"results/OLS_models.png",sep=""), units = "in", width = 10, height = 5, res = 300)
ggplot(model.summary.realm[
  -which(
    model.summary.realm$REALM == "Panamanian" |
      model.summary.realm$REALM == "Oceanian" |
      model.summary.realm$REALM == "Madagascan"),], 
  aes(x = var, y = Estimate, 
      color = realm,
      alpha = sig)) +
  geom_linerange(aes(ymin=Lower_IC, 
                     ymax=Higher_IC),
                 position=position_dodge(0.5)) +
  geom_point(data = subset(model.summary.2, realm == "Global"), 
             aes(x = var, y = Estimate), 
             size = 4, color = "black") +
  geom_linerange(data = subset(model.summary.2, realm == "Global"),
                 aes(ymin=Lower_IC, 
                     ymax=Higher_IC), color = "black") +
  geom_point(size = 2,
             position=position_dodge(0.5)) +
  scale_color_manual(values = col_realms)+
  geom_hline(yintercept=0,linetype = "dashed") +
  
  xlab("") + coord_flip() + 
  scale_radius()+
  theme_bw() +
  #theme(text = element_text(size=20))+
  facet_grid(model ~ taxa, scales = "free")

dev.off()
```

# SAR models
## Define define a connectivity (neighbourhood) matrix
```{r}

ncores <- detectCores()
cl <- makeCluster(ncores)
registerDoParallel(cl)

start_time <- Sys.time()

nlw <- 
  foreach(i=1:length(taxa), .packages=c('spdep')) %dopar% { 
    # define connectivity matrix (0/1)
    nbdist<-dnearneigh(x=as.matrix(dataset.group[[i]][,1:2]), d1=0, d2=5000) 
    
    # compute the Euclidean distance between neighbouring sites
    # neigh.dist<-nbdists(nbdist, 
    #                     as.matrix(dataset.group[[i]][,1:2]), longlat=F)
    # # compute the inverse distance weigthed matrix
    # inverse<-lapply(neigh.dist, function(x) (1/(x^2))) 
    # coding style W = row standardised
    nb2listw(neighbours=nbdist, 
             # glist=inverse, 
             style="W", 
             zero.policy=FALSE) 
  }

names(nlw) <- taxa

# stop the cluster
stopCluster(cl)

end_time <- Sys.time()
end_time-start_time
# takes 5 minutes

```

## Global models
```{r warning= F}

#load("Code/sarmodels.rds")

ncores <- detectCores()
cl <- makeCluster(ncores)
registerDoParallel(cl)

start_time <- Sys.time()

sar.models <- 
  foreach(i=1:length(taxa), .packages=c('spdep')) %dopar% { 
    
    models.list <- list()
    
    model.age <- spatialreg::errorsarlm(Age ~ 
                                          Stab,
                                        data = dataset.group[[i]],
                                        listw = nlw[[i]])
    
    model.div <- spatialreg::errorsarlm(Div ~ 
                                          Age+
                                          Stab,
                                        data = dataset.group[[i]],
                                        listw = nlw[[i]])
    
    model.div2 <- spatialreg::errorsarlm(Div ~ 
                                           Age+
                                           Stab+
                                           Prod,
                                         data = dataset.group[[i]],
                                         listw = nlw[[i]])
    
    model.FD <- spatialreg::errorsarlm(FD ~ 
                                         Div+
                                         Age+
                                         Stab+
                                         Prod,
                                       data = dataset.group[[i]], 
                           listw = nlw[[i]])
    
    models.list$div <- model.div
    models.list$div2 <- model.div2
    models.list$age <- model.age
    models.list$FD <- model.FD
    
    models.list
    
  }

# stop the cluster
stopCluster(cl)

end_time <- Sys.time()
end_time-start_time
# takes 2.37 hours

# saveRDS(sar.models, "sarmodels.rds")
# sar.models <- readRDS("sarmodels.rds")


```

### Get results from global models
```{r echo=F, eval=FALSE}

model.summary.sar <- list()

for(i in 1:length(taxa)){
  
  model.div <- summary(sar.models[[i]]$div)
  model.div.vars <- rownames(model.div$Coef)[-1]
  model.div.summary <- model.div$Coef[-1,]
  
  model.FD <- summary(sar.models[[i]]$FD)
  model.FD.vars <- rownames(model.FD$Coef)[-1]
  model.FD.summary <- model.FD$Coef[-1,]
  
  model.summary.sar_ <- data.frame(model = "FD",
                     data.frame(
                       var = model.FD.vars,
                       model.FD.summary),
                     confint(sar.models[[i]]$FD)[-1:-2,])
  
  names(model.summary.sar_)[7] <- "Lower_IC"
  names(model.summary.sar_)[8] <- "Higher_IC"
  
  model.summary.sar_$sig <- ifelse(model.summary.sar_$Pr...z..<0.05,1,0)
  
  model.summary.sar[[i]] <- model.summary.sar_
}
names(model.summary.sar) <- taxa

model.summary.sar.2 <- rbindlist(model.summary.sar,idcol = T)
names(model.summary.sar.2)[1] <- "taxa"
```

### Get residuals from global models
```{r}
res.sar.FD <- list()
res.lm.FD <- list()

for(i in 1:length(taxa)){
  res.sar.FD[[i]] <- residuals(sar.models[[i]]$FD)
  res.lm.FD[[i]] <- residuals(lm.models[[i]]$FD)
}
```

## Run Regional SAR models
```{r}

model.summary.sar.realm <- list()

res.sar.FD.realm <- list()

## Models by realm

for(i in 1:length(taxa)){ 
  cat("/r","Taxa", taxa[i], '-', i, 'from', length(taxa))
  
  cl <- makeCluster(8)
  registerDoParallel(cl)
  
  sar.models.realm <- 
    foreach(j=1:NRealm, .packages=c('spdep')) %dopar% { 
      
      models.list <- list()
      
      # get regional dataset
      datatmp <- reg_dataset[[i]][[j]]
      # Convert to SpatialPointsdataframe
      coordinates(datatmp) <- ~x+y 
      # define connectivity matrix (0/1)
      nbdist<-dnearneigh(x=datatmp, d1=0, d2=5000) 
      # compute the Euclidean distance between neighbouring sites
      # neigh.dist<-nbdists(nbdist, datatmp, longlat=F) 
      # 
      # inverse<-lapply(neigh.dist, function(x) (1/(x^2)))
    
      # coding style W 
      nlw2<-nb2listw(neighbours=nbdist, 
                     # glist=inverse, 
                     style="W", zero.policy=FALSE) 
      
      
      models.list <- list()
      
      model.age <- spatialreg::errorsarlm(Age ~ 
                                            Stab,
                                          data = datatmp,
                                          listw = nlw2)
      
      model.div <- spatialreg::errorsarlm(Div ~ 
                                            Age+
                                            Stab,
                                          data = datatmp,
                                          listw = nlw2)
      
      model.div2 <- spatialreg::errorsarlm(Div ~ 
                                             Age+
                                             Stab+
                                             Prod,
                                           data = datatmp,
                                           listw = nlw2)
      
      model.FD <- spatialreg::errorsarlm(FD ~ 
                                           Div+
                                           Age+
                                           Stab+
                                           Prod,
                                         data = datatmp, 
                                         listw = nlw2)
      
      models.list$div <- model.div
      models.list$div2 <- model.div2
      models.list$age <- model.age
      models.list$FD <- model.FD
      
      models.list
      
    }
  # stop the cluster
  stopCluster(cl)
  
  res.sar.rich.realm_ <- list()
  res.sar.div.realm_ <- list()
  res.sar.FD.realm_ <- list()
  
  model.summary.sar.realm_ <- list()
  
  for(j in 1:NRealm){ 
    
    res.sar.FD.realm_[[j]] <- residuals(sar.models.realm[[j]]$FD)
    
    model.FD.2 <- summary(sar.models.realm[[j]]$FD)
    model.FD.summary <- model.FD.2$Coef[-1,]
    
    model.summary.sar_ <-             data.frame(model = "FD",
                       data.frame(
                         var = rownames(model.FD.summary),
                         model.FD.summary),
                       confint(sar.models.realm[[j]]$FD)[-1:-2,])
    
    names(model.summary.sar_)[7] <- "Lower_IC"
    names(model.summary.sar_)[8] <- "Higher_IC"
    
    model.summary.sar_$sig <- ifelse(model.summary.sar_$Pr...z..<0.05,1,0)
    
    model.summary.sar.realm_[[j]] <- model.summary.sar_
    
  }
  names(model.summary.sar.realm_) <- Realms
  
  model.summary.sar.realm[[i]] <- rbindlist(model.summary.sar.realm_,idcol = T)
  
  names(res.sar.FD.realm_) <- Realms
  
  res.sar.FD.realm[[i]] <- res.sar.FD.realm_
  
  rm(sar.models.realm)
}

names(model.summary.sar.realm) <- taxa

model.summary.sar.realm <- rbindlist(model.summary.sar.realm,idcol = T)
names(model.summary.sar.realm)[1:2] <- c("taxa", "realm")

names(res.sar.FD.realm) <- taxa


```


## Plot model coeffs SAR models
```{r warning= F}
col_realms <- c("darkgreen", # dark green
                "#0072B2", # blue
                "#00BE67", # green
                "#56B4E9", # skyblue
                "#E69F00", # orange
                "#D55E00") # vermillion
names(col_realms) <- Realms

model.summary.sar.realm$taxa <- factor(model.summary.sar.realm$taxa, 
                                        levels=c('birds','mammals','amphibians','reptiles'))
model.summary.sar.realm$var <- factor(model.summary.sar.realm$var, 
                                        levels=c('Stab','Age','Div','Prod'))
model.summary.sar.realm$realm <- factor(model.summary.sar.realm$realm, 
                                        levels=c('Afrotropics',
                                                 'Neotropics',
                                                 'Palearctic',
                                                 'Nearctic',
                                                 'IndoMalay',
                                                 'Australasia'))

model.summary.sar.2$taxa <- factor(model.summary.sar.2$taxa, 
                                        levels=c('birds','mammals','amphibians','reptiles'))
model.summary.sar.2$var <- factor(model.summary.sar.2$var, 
                                        levels=c('Stab','Age','Div','Prod'))
model.summary.sar.2$realm <- factor(model.summary.sar.2$realm, 
                                        levels=c('Afrotropics',
                                                 'Neotropics',
                                                 'Palearctic',
                                                 'Nearctic',
                                                 'IndoMalay',
                                                 'Australasia'))

pdf(paste(savedatafolder,"results/SAR_models.pdf",sep=""), width = 9, height = 3)

ggplot(data = model.summary.sar.realm, 
  aes(x = var, y = Estimate, color = realm, alpha = as.factor(sig))) +
  geom_hline(yintercept=0, linetype = "dashed") +
  geom_linerange(data = model.summary.sar.2,
                 aes(ymin=Lower_IC, 
                     ymax=Higher_IC), 
                 color = "black") +
  geom_point(data = model.summary.sar.2, 
             aes(x = var, y = Estimate), 
             size = 4, color = "black") +
  geom_linerange(aes(ymin=Lower_IC, 
                     ymax=Higher_IC),
                 position=position_dodge(0.5)) +
  geom_point(size = 2,
             position=position_dodge(0.5)) +
  scale_color_manual(values = col_realms)+
  ylab("Std. Coeff.") + xlab("") + coord_flip() + 
  scale_radius()+
  theme_bw() +
  facet_grid(. ~ taxa, scales = "free")

dev.off()

```

## Correlogram global models
```{r}

FD.sar.obs.cor <- list()
FD.sar.res.cor <- list()
FD.lm.res.cor <- list()

for(i in 1:length(taxa)){ cat("Calculating correlograms for", taxa[i], "\n")
  
  res.sar.FD <- sar.models[[i]]$FD$residuals
  FD.index = sample(1:length(res.sar.FD),1000,replace = F)
  
  # Observed FD
  FD.sar.obs.cor[[i]]<-as.data.frame(
    pgirmess::correlog(coords=XY[[i]][FD.index,], 
                       z=dataset.group[[i]]$FD[FD.index], 
                       method="Moran"))
  
  # Res sar model
  FD.sar.res.cor[[i]]<-as.data.frame(
    pgirmess::correlog(coords=XY[[i]][FD.index,], 
                       z=res.sar.FD[FD.index], 
                       method="Moran"))
  
  # Res lm model
  res.lm.FD <- lm.models[[i]]$FD$residuals
  FD.index = sample(1:length(res.lm.FD),1000,replace = F)
  
  FD.lm.res.cor[[i]]<-as.data.frame(
    pgirmess::correlog(coords=XY[[i]][FD.index,], 
                       z=res.lm.FD[FD.index], 
                       method="Moran"))
}

# plot
pdf("results/correlograms.pdf", 10,2.5)
par(mfrow=c(1,4))
for(i in 1:length(taxa)){ 
  {plot(main = taxa[i],
        col = "white",
        coef~dist.class, 
        xlab="Geographic distance", 
        ylab="Moran's I", 
        pch=1, bty="l", cex.lab=1.2, 
        xlim=c(0,20000), ylim=c(-1,1), 
        data=FD.sar.res.cor[[i]])
    
    abline (h=0, lty=2, col="black")
    
    # add connected line segments to Morans'I of the raw value
    lines(FD.sar.obs.cor[[i]]$dist.class, 
          FD.sar.obs.cor[[i]]$coef, 
          col = "black", lwd=2) 
    
    # add connected line segments to Morans'I of lm model residuals
    lines(FD.lm.res.cor[[i]]$dist.class, 
          FD.lm.res.cor[[i]]$coef, 
          col = "red", lwd=2) 
    
    # add connected line segments to Morans'I of SAR model residuals
    lines(FD.sar.res.cor[[i]]$dist.class, 
          FD.sar.res.cor[[i]]$coef, 
          col = "blue", lwd=2) 
    
    legend((FD.sar.obs.cor[[i]][14,1])/4, 1, 
           c("Observed values", 
             "LM residuals",
             "SAR residuals"), 
           y.intersp = 1,
           pch="_", bty="n", cex=1, 
           col = c("black","red","blue"))
  }
}
dev.off()

```

# Piecewise sem lm
```{r echo=F, eval=FALSE}
options(na.action("na.omit"))

sem.lm <- list()
dataset.group <- list()
res.sem <- list()

for(i in 1:length(taxa)){
  
  cat("Running models for", taxa[i])
  
  model.FD <- lm(FD ~ 
                    Div+
                    Age+
                    Prod+
                    Stab,
                  data = dataset.group[[i]], 
                  na.action=na.fail)
  
  model.div <- lm(Div ~ 
                    Stab+
                    Prod,
                  data = dataset.group[[i]], 
                  na.action=na.fail)
  
  model.age <- lm(Age ~ 
                    Stab+
                    Prod,
                  data = dataset.group[[i]], 
                  na.action=na.fail)
  
  sem.lm.tmp <- psem(
    model.FD,
    model.div,
    model.age)
  
  sem.lm[[i]] <- summary(sem.lm.tmp, .progressBar = F)
  
  res.sem[[i]] <- residuals(sem.lm.tmp)
  
}


```

# Piecewise sem sar
## Global models
```{r echo=F, eval=FALSE}
options(na.action("na.omit"))

start_time <- Sys.time()

# ncores <- detectCores()
# cl <- makeCluster(ncores)
# registerDoParallel(cl)

sar.sem <- list()
# sar.sem <-  foreach(i=1:length(taxa),
#                     .packages=c('spdep','spatialreg','piecewiseSEM')) %dopar% {

for(i in 1:length(taxa)){    cat("\r",taxa[i])
  
  data.tmp <- dataset.group[[i]]
  neigh <- nlw[[i]]
  
  model.age <- spatialreg::errorsarlm(Age ~ 
                                        Stab,
                                      data = data.tmp,
                                      listw = neigh)
  
  model.div <- spatialreg::errorsarlm(Div ~ 
                                        Age+
                                        Stab,
                                      data = data.tmp,
                                      listw = neigh)
  
  model.div2 <- spatialreg::errorsarlm(Div ~ 
                                         Age+
                                         Stab+
                                         Prod,
                                       data = data.tmp,
                                       listw = neigh)
  
  model.FD <- spatialreg::errorsarlm(FD ~ 
                                       Div+
                                       Age+
                                       Stab+
                                       Prod,
                                     data = data.tmp,
                                     listw = neigh)
  
  cat("\r",taxa[i], "SEM_model")
  
  sem.sar.tmp <- psem(
    model.FD,
    model.div,
    model.age)
  
  cat("\r",taxa[i], "SEM2_model")
  sem.sar.tmp2 <- psem(
    model.FD,
    model.div2,
    model.age)
  
  models.sarsem <- list(model1 = list(summary=list(),
                                        residuals=list()), 
                          model2 = list(list(summary=list(),
                                        residuals=list())))
  
  cat("\r",taxa[i], "SEM_model")
  models.sarsem$model1$summary <- lapply(list(model.FD,
                                              model.div,
                                              model.age), 
                                         function(x)
                                           summary(x, Nagelkerke=T))
  models.sarsem$model1$residuals <- residuals(sem.sar.tmp)
  models.sarsem$model1$psem <- summary(sem.sar.tmp)
  
  cat("\r",taxa[i], "SEM_model 2")
  models.sarsem$model2$summary <- lapply(list(model.FD,
                                              model.div2,
                                              model.age), 
                                         function(x)
                                           summary(x, Nagelkerke=T))
  models.sarsem$model2$residuals <- residuals(sem.sar.tmp2)
  models.sarsem$model2$psem <- summary(sem.sar.tmp2)
  
  # models.sarsem
  sar.sem[[i]] <- models.sarsem
  
}


# stop the cluster
stopCluster(cl)

end_time <- Sys.time()
end_time-start_time
# Takes 10 hours

names(sar.sem) <- taxa

# saveRDS(sar.sem, "sarsem.rds")

```

### Get results from global models
```{r echo=F, eval=FALSE}
sarsem_res <- list()
sarsem_model1  <- list()
sarsem_model2  <- list()

for(i in 1:length(taxa)){
  
  sarsem_model1[[1]] <- sar.sem[[i]]$model1$psem$coefficients
  
  sarsem_model1[[2]] <- rbind(data.frame(Y = "FD", 
                                         R2 = round(summary(sar.models[[i]]$FD, Nagelkerke = T)$NK,2)),
                              data.frame(Y = "Div",  
                                         R2 = round(summary(sar.models[[i]]$div, Nagelkerke = T)$NK,2)),
                              data.frame(Y = "Age", 
                                         R2 = round(summary(sar.models[[i]]$age, Nagelkerke = T)$NK,2)))
  
  sarsem_model2[[1]] <- sar.sem[[i]]$model2$psem$coefficients
  
  sarsem_model2[[2]] <- rbind(data.frame(Y = "FD", 
                                         R2 = round(summary(sar.models[[i]]$FD, Nagelkerke = T)$NK,2)),
                              data.frame(Y = "Div",  
                                         R2 = round(summary(sar.models[[i]]$div2, Nagelkerke = T)$NK,2)),
                              data.frame(Y = "Age", 
                                         R2 = round(summary(sar.models[[i]]$age, Nagelkerke = T)$NK,2)))
  
  sarsem_res[[i]] <- list(sarsem_model1, sarsem_model2)
  names(sarsem_res[[i]]) <- c("model1", "model2")
  
}
names(sarsem_res) <- taxa

```

### Plot global SEMs
```{r echo=F, eval=FALSE}
# Transforming a model summary into a graph description
# must use DiagrammeR v0.6

for(i in 1:length(taxa)){
  node_set <- data.frame(
    nodes = c("FD", "Age", "Div", "Prod", "Stab"),
    shape = c("box", "box", "box", "box", "box")
  )
  
  edge_set  <- data.frame(
    edge_from = sarsem_res[[i]]$model2[[1]]$Predictor,
    edge_to = sarsem_res[[i]]$model2[[1]]$Response,
    style = ifelse(sarsem_res[[i]]$model2[[1]]$P.Value<.05,"solid","dashed"),
    label = round(sarsem_res[[i]]$model2[[1]]$Estimate,2),
    penwidth = abs(sarsem_res[[i]]$model2[[1]]$Estimate)*20,
    color = ifelse(sarsem_res[[i]]$model2[[1]]$Estimate<0,"red","blue")
  )
  
  # Combine edges and nodes
  my_graph <- graphviz_graph(
    nodes_df = node_set,
    edges_df = edge_set,
    edge_attrs = c("arrowsize = .5"))
  
  # cat(my_graph$dot_code)
  
  graphviz_render(my_graph)
  
  graphviz_render(my_graph) %>%
    export_svg %>% charToRaw %>% rsvg_pdf(paste("results/SEM_global_",taxa[i],".pdf",sep = ""))
}


```



# SAVE
```{r echo=F, eval=FALSE}
# save.image("FD_global_spatial_analyses.RData")
```
