---
title: "FD project (reptiles)"
author: "Brunno F Oliveira^1^; Brett R Scheffers^2^;"
date: "May 9, 2017 **1** Department of Wildlife Ecology and Conservation, University of Florida/IFAS, Gainesville, FL 32611, USA. **Corresponding author:** (brunno.oliveira@me.com)"
output:
  pdf_document: default
theme: united
toc: yes
---

*** 

\newpage



# Packages versions:
```{r info,message=FALSE, echo=F}
info <- sessionInfo()
```

We used `r info[1]$R$ver` and the following packages:

```{r packages, echo=FALSE, message=FALSE, warning=FALSE, include=T}
rm(list=ls())
gc()

list.of.packages <- c("picante","ape","raster","maptools","rgeos","stringr","rgdal","ggplot2","plyr","grid","gridExtra","multcompView","caper","geiger","phytools","knitr","maps","parallel","doParallel","apTreeshape","data.table","ppcor", "relaimpo","FD","knitr","PhyloMeasures","PerformanceAnalytics","pbapply","dplyr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

sapply(list.of.packages, require, character.only = TRUE)
```


```{r "setup"}
#WHERE ARE YOU?

savedatafolder <- "C:/Users/boliveir/Dropbox (UFL)/FD_causes/" # address to folder to save results
externHD <- 'G:/GIS/' # address to external HD
setwd("C:/Users/boliveir/Dropbox (UFL)/FD_causes")
knitr::opts_knit$set(root.dir = "C:/Users/boliveir/Dropbox (UFL)/FD_causes")

#getwd()
# list.files()

# function to cal MPD for a list of species
# cophe_phyd = cophenetic distance matrix
# splist= sp list
mpd_b <- function(cophe_phyd, splist){
  dist_trx <- cophe_phyd[splist, splist]
  MPD_ <- mean(dist_trx[lower.tri(dist_trx)])
  return(MPD_)
}

# load function to calculate fdips in the loop
source("fdisp_loop.R")


```

## Land data
```{r , echo=F}
land <- rgdal::readOGR("G:/GIS/Shp files/ne_50m_land/ne_50m_land_no_artic.shp")
crs(land) <-"+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 "
land <- spTransform(land, CRS("+proj=cea +datum=WGS84"))

  
# REFERENCE PROJECTION MAP
mapa <- raster(xmn = -20592508, xmx = 20588492, ymn = -5743602, ymx = 6573398,
              crs = CRS("+proj=cea +datum=WGS84"))
res(mapa) <- 111000
```

# Load species pres/abs:
## Species occurrences
```{r species occr, eval=T, message=TRUE}
# Are you running the code for the first time?

answer <- "yes"

if(answer=="no"){
  load(paste(savedatafolder,"code/reptiles",'_data.RData',sep=""))
  #load(paste(savedatafolder,"code/reptiles",'_data_cluster.RData',sep=""))
}
if(answer=="yes"){
  
occr <- read.csv(paste(
  savedatafolder,"/Data/reptiles_occr_equal_area_30land.csv",sep=""))
XY <- occr[,1:2]
names(XY) <- c("x","y")
coordinates(XY)<-~x+y
occr <- occr[,-1:-2]
amph_sp_list <- colnames(occr)
if(anyDuplicated(amph_sp_list)>0){
  occr <- occr[,-which(duplicated(amph_sp_list))]
}

# Check
#Species richness
{test <- data.frame(XY,rowSums(occr)) 
colnames(test)[1:2] <- c("x","y")
coordinates(test) <- ~x+y
gridded(test) <- TRUE
test <- raster(test)}
{plot(main="Richness", test)
map(land,add=T,cex=.5)} 

spp <- colnames(occr)
spp <- data.frame(gsub("[.]","_",spp))
names(spp) <- "sp"
}

colnames(occr) <- spp$sp
```

The original species occurrence dataset has `r nrow(spp)` species

## Traits
Using traits of lizards
Shai Meiri. 2018. Traits of lizards of the world: Variation around a successful evolutionary design. GEB.https://doi.org/10.1111/geb.12773 
```{r open_traits, eval=F}
trait <- data.frame(read.csv(paste(sep="",savedatafolder,"Data/Meiri_-_Lizard_data.csv"), header=TRUE))

names(trait) 

# Get traits of interest
trait <- data.frame(trait[,c(1,11,13,14,18,20,22,24,26,31,36,37)]) 

names(trait)[1] <- "species"

trait_data <- data.frame(gsub(" ","_",trait$species))
rownames(trait) <- trait_data$gsub......._...trait.species.

trait <- trait[,-1]
trait$maximum.SVL <- log(trait$maximum.SVL+1)
trait$hatchling.neonate.SVL <- log(trait$hatchling.neonate.SVL+1)
trait$largest.mean.clutch.size <- log(trait$largest.mean.clutch.size+1)
trait$minimum.mean.Tb <- log(trait$minimum.mean.Tb+1)
trait$maximum.mean.Tb <- log(trait$maximum.mean.Tb+1)


```

The original trait dataset has `r nrow(trait)` species.  

## Phylogeny
Using Jetz and Pyron 2017
```{r open_traits, eval=F}
### OPEN LARGE DATASET
tre <- read.tree(paste(sep="",savedatafolder,"Data/squam_shl_new_Consensus_9755.tre"))
tredata <- data.frame(tre$tip.label)
```
The original tree has `r nrow(tredata)` species.  



## Get species in common among the three datasets
```{r count_traits, eval=F}
names(spp) <- names(trait_data) <- names(tredata) <- "sps"

lista <- merge(data.frame(spp),data.frame(trait_data),by="sps")
lista <-as.vector(lista[,1])
lista <- unique(lista)
```

There are `r length(lista)` species in common between the three datasets (`(length(lista)*100)/nrow(spp)`% of all species present in the IUCN dataset). We use this species' list to subset the original species and trait data. This dataset will be used in the following analyses.

## Subseting species from datasets
```{r eval=F}
# 1 - DISTRIBUTIONAL DATA
occr <- subset(occr, select=lista) # Subset the species beloguing to both datasets
Rich <- rowSums(occr) # Get richness for every cell

# subset communities with less than 4 species
rem <- which(Rich < 4)
occr <- occr[-rem,]
Rich <- Rich[-rem]
XY <- as.data.frame(XY)[-rem,]
occr <- occr[,-which(colSums(occr)==0)] # Remove species that does not occur in any community
# new list of species
lista <- colnames(occr)

gc() # free up more memmory

# 2 - PHYLOGENY
phylo_nqro <- drop.tip(tre,lista) #Remove as espécies que quero e ficam as que não quero
tip_nqro<-as.vector(phylo_nqro$tip.label) #Lista das sps que nao quero
tre<-drop.tip(tre, tip_nqro) #Eliminar as sp que nao quero e ficam as que quero

rm(tip_nqro,phylo_nqro)

# 3 - TRAIT DATA
trait <- trait[rownames(trait) %in% lista,]

# 4 - REARRANGE DATA
# The species data in the data frames containing the distribution and trait measurements 
# must be in the same row order as the names of the species in the tree object (mytree$tip.label). 
occr <- occr[,match(tre$tip.label,colnames(occr))]
trait <- trait[match(tre$tip.label,rownames(trait)),]
# check whether our community, tree and traits are in the same order.
all.equal(colnames(occr), tre$tip.label)
all.equal(colnames(occr), rownames(trait))

rm(trait_data,tredata,spp)

gc() # free up more memmory
```
After removing communities with less than 4 species, our final data contained `r length(lista)` species.

#Load environmental variables 
```{r}
# create raster with the same extent of `Richness_Raster` (~1 degree)
r <- mapa

res(r)
```

## Climate change velocity (Sandel et al. 2011)
Downloaded from [Dryad](http://datadryad.org/resource/doi:10.5061/dryad.b13j1).
Citation: Sandel, B. et al. (2011) The influence of Late Quaternary climate-change velocity on species endemism. Science.
```{r get_velocity, eval=F}
velo.r <- raster(paste(sep="",externHD,'#Environment/climate_velocity/Velocity.tif'))
velo.r <- projectRaster(velo.r, crs=crs(mapa))

velo.r <- resample(velo.r, r, method= "bilinear")
velocity <- extract(velo.r,XY,method='simple')

rm(velo.r)
```

## Climate variance over time (Hannah)
```{r get_velocity, eval=F}
# Stability over time metric
climate.stab.r <- raster(paste(sep="",externHD,'#Environment/Climate_Stability_Hannah_Owens/climateStability.asc'))
crs(climate.stab.r) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 "
climate.stab.r <- projectRaster(climate.stab.r, crs=crs(mapa))

climate.stab.r <- resample(climate.stab.r, r, method= "bilinear")
climate.stab <- extract(climate.stab.r,XY,method='simple')

rm(climate.stab.r)
```

## Climate distance and novelty (Fitzpatrick et al. 2018)
Citation: Fitzpatrick et al. (2018) How will climate novelty influence ecological forecasts? Using the Quaternary to assess future reliability. Glob. Change Bio.
```{r get_clim_dist, eval=F}
# Temperature distance
tempdist.r <- raster(paste(sep="",externHD,'#Environment/Climate_distance_novelty/climate_distance/climate_distance_pairwiseavg_halfdegree.temp.tif'))
tempdist.r <- projectRaster(tempdist.r, crs=crs(mapa))

tempdist.r <- resample(tempdist.r, r, method= "bilinear")
tempdist <- extract(tempdist.r,XY,method='simple')

rm(tempdist.r)

# Precip distance
precipdist.r <- raster(paste(sep="",externHD,'#Environment/Climate_distance_novelty/climate_distance/climate_distance_pairwiseavg_halfdegree.precip.tif'))
precipdist.r <- projectRaster(precipdist.r, crs=crs(mapa))

precipdist.r <- resample(precipdist.r, r, method= "bilinear")
precipdist <- extract(precipdist.r,XY,method='simple')

rm(precipdist.r)

# Temperature novelty
tempnov.r <- raster(paste(sep="",externHD,'#Environment/Climate_distance_novelty/climate_novelty/climate_novelty_pairwiseavg_halfdegree.temp.tif'))
tempnov.r <- projectRaster(tempnov.r, crs=crs(mapa))

tempnov.r <- resample(tempnov.r, r, method= "bilinear")
tempnov <- extract(tempnov.r,XY,method='simple')

rm(tempnov.r)

# Precip novelty
precipnov.r <- raster(paste(sep="",externHD,'#Environment/Climate_distance_novelty/climate_novelty/climate_novelty_pairwiseavg_halfdegree.precip.tif'))
precipnov.r <- projectRaster(precipnov.r, crs=crs(mapa))

precipnov.r <- resample(precipnov.r, r, method= "bilinear")
precipnov <- extract(precipnov.r,XY,method='simple')

rm(precipnov.r)

```

## Canopy height and biomass at 1km spatial resolution.
Downloaded from [NASA](http://lidarradar.jpl.nasa.gov/).  
Citation: Simard, M. et al. (2011), "Mapping forest canopy height globally with spaceborne lidar", Journal of Geophysical Research, VOL. 116, G04021. 
```{r get_tree_heigh, eval=F}
veg.r <- raster(paste(sep="",externHD,'#Environment/NASA veg height/asc/veg-height_1dg-equal.asc'))

veg <- extract(veg.r,XY,method='simple')

rm(veg.r)
```

## Current climate conditions (Hijmans et al. 2005)  
Downloaded from [Worldclim project](http://www.worldclim.org) at the resolution of 10 minutes of a degree ~ 1 degree
Citation: Hijmans, R.J., et al. (2005) Very high resolution interpolated climate surfaces for global land areas. International Journal of Climatology
```{r get_current_climate, eval=F}
bio <- raster::getData("worldclim", var = "bio", res = 10)
bio <- projectRaster(bio, crs=crs(mapa))
  
mean.temp.r <- resample(bio$bio1,r, method= "bilinear") # Annual mean temperature
max.temp.r <-  resample(bio$bio5,r, method= "bilinear") # Max Temperature of Warmest Month
min.temp.r <-  resample(bio$bio6,r, method= "bilinear") # Min Temperature of Warmest Month
sea.temp.r <- resample(bio$bio4,r, method= "bilinear") # temperature seasonality
diu.temp.r <- resample(bio$bio2,r, method= "bilinear") # diurnal temperature range

mean.temp <- extract(mean.temp.r,XY,method='simple')
max.temp <- extract(max.temp.r,XY,method='simple')
min.temp <- extract(min.temp.r,XY,method='simple')
sea.temp <- extract(sea.temp.r,XY,method='simple')
diu.temp <- extract(diu.temp.r,XY,method='simple')

mean.prec.r <- resample(bio$bio12,r, method= "bilinear") # Annual preciptation
sea.prec.r <- resample(bio$bio15,r, method= "bilinear") # Precipitation seasonality

mean.prec <- extract(mean.prec.r,XY,method='simple')
sea.prec <- extract(sea.prec.r,XY,method='simple')

rm(bio, mean.temp.r,max.temp.r,min.temp.r,sea.temp.r,diu.temp.r,
   mean.prec.r,sea.prec.r)
```


## NPP
Downloaded from [NASA](https://lpdaac.usgs.gov/dataset_discovery/modis/modis_products_table/mod17a3)
Citation: Running, S. W. et al. 2004. A Continuous Satellite-Derived Measure of Global Terrestrial Primary Production. - Bioscience 54: 547-560.

Valid range of values is 0-65500. Values > 65500 represent other categories of land
see https://lpdaac.usgs.gov/dataset_discovery/modis/modis_products_table/mod17a3
I found that there are much cells with values bellow 65500 that are unlikely to be correct values. For instance, see Fig 2 in S?mov?, I. & Storch, D. (2017) The enigma of terrestrial primary productivity: measurements, models, scales and the diversity-productivity relationship. Ecography, 40, 239-252. Here, the max NPP value is around 2000. Thus, I modified the original NPP raster by 1) giving NA to cells > 65500, and then 2) giving 0 to cells > 30000.

```{r get_tree_heigh, eval=F}
# productivity
NPP <- raster(paste(sep="",externHD,'#Environment/NPP_GPP_means_2000_2012/MOD17A3_Science_NPP_mean_00_12_equal-area_NULL_30000.tif'))

NPP <- extract(NPP,XY,method='simple')
```


## Evelation data at 1 degree.
Downloaded from [Worldclim project](http://www.worldclim.org) at the resolution of 10 minutes of a degree ~ 1 degree. Aggregated from SRTM 90 m resolution
```{r get_elevation, eval=F}
#  topographic indeces 
elevation <- raster::getData('worldclim', var='alt', res = 2.5)
elevation <- projectRaster(elevation, crs=crs(mapa))
elev <- extract(elevation,XY,method='simple')

terrain_indices <- terrain(elevation, opt=c('slope','aspect', 'TPI', 'TRI', 'roughness'))
terrain_indices <- extract(terrain_indices,XY,method='simple')

terrain_indices <- data.frame(elev, terrain_indices)

rm(elevation)
```


## Realms
```{r get_realm, warning= F, eval=F}
Reanames <- readShapePoly("G:/GIS/CMEC regions & realms/newRealms.shp")
  crs(Reanames) <-"+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 "
  extent(mapa) <- extent(Reanames)
  Reanames<-data.frame(Reanames$Realm)
  names(Reanames)<-'Realm'
  Reanames<-cbind(Realms=rownames(Reanames),Reanames)
  #2) Extract values
  Realm <- readShapePoly("G:/GIS/CMEC regions & realms/newRealms.shp")
  crs(Realm) <-"+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 "
  realmnames<-sort(Realm$fullupgmar) # put alphabetic order because rasterize will put number in acordance with alphabetic order
  extent(mapa) <- extent(Realm)
  Realm <- rasterize(Realm,mapa,'fullupgmar')
  Realm <- projectRaster(Realm, crs="+proj=cea +datum=WGS84")
  Realm <- as.vector(raster::extract(Realm,XY,method='simple'))
  
  #Some cells comprise more than one Realm and the extract function give a medium value for that cell. 
  #This value have no meaning as we need the value of the Realm for each cell.
  check.integer <- function(N){ # function for check if number is integer
    !length(grep("[^[:digit:]]", as.character(N)))
  }
  
  for (i in seq_along(Realm)){ # put NA in non-integer numbers
    if (check.integer(Realm[i]))
    {Realm[i]<-Realm[i]
    }else{Realm[i]<-NA}
  }
  
  #Dont know why...but some codes come with whitespaces -To see use: unique(Realm)
  trim <- function( x ) { #Function to remove whitespaces
    gsub("(^[[:space:]]+|[[:space:]]+$)", "", x)
  }
  Realm<-trim(Realm)
  
  #Give names for Realm code...
  Realm <- as.character(Reanames$Realm)[ match(Realm, c(1:11))]
  
  ### Solve problem with NAs between realms
  Realm2 <- readShapePoly("G:/GIS/CMEC regions & realms/newRealms.shp")
  crs(Realm2) <-"+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 "
  Realm2 <- spTransform(Realm2, CRS("+proj=cea +datum=WGS84"))
  XY <- SpatialPoints(XY)
  crs(XY) <-"+proj=cea +datum=WGS84"
  Realm2<-sp::over(XY,Realm2)
  Realm2<-as.vector(Realm2$Realm)
  
  #Give names for NAs based on the previous 
  Realm3<-NA
  for(i in 1:length(Realm2)){
    if(is.na(Realm[i])){
      Realm3[i]<-Realm2[i]
    }else{
      Realm3[i]<-Realm[i]
    } 
  }
  
  Realm<-Realm3
  rm(Realm2,Realm3)
```


## WWF Biomes
```{r biomes}
biome <- readShapePoly(paste(sep="",externHD,"WWF_biomas_ecoregions/wwf_terr_ecos.shp"))
crs(biome) <-"+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 "
biome <- spTransform(biome, crs(r))
biome <- rasterize(biome,r,'BIOME')
biome <- as.vector(extract(biome,XY,method='simple',fun=min(x)))

biome<-gdata::trim(biome)
biome<-round(as.numeric(biome),0)
biome[which(biome>14)] <- NA

# biomes names
bionames <- c(
"Tropical & Subtropical Moist Broadleaf Forests",
"Tropical & Subtropical Dry Broadleaf Forests", 
"Tropical & Subtropical Coniferous Forests",
"Temperate Broadleaf & Mixed Forests",
"Temperate Conifer Forests",
"Boreal Forests/Taiga",
"Tropical & Subtropical Grasslands, Savannas & Shrublands",
"Temperate Grasslands, Savannas & Shrublands",
"Flooded Grasslands & Savannas",
"Montane Grasslands & Shrublands",
"Tundra",
"Mediterranean Forests, Woodlands & Scrub",
"Deserts & Xeric Shrublands",
"Mangroves")
```

## Biorealm
```{r get_tree_density}
Biorealm <- paste(Realm, biome,sep = "_")
```


## Environmental variables
```{r env_vars}
env.vars <- data.frame(velocity,climate.stab,tempnov,tempdist,veg,
                       mean.temp,mean.prec,elev,terrain_indices,
                       NPP,Realm,Biorealm,biome)
```

# COMMUNITY METRICS
```{r community_metrics, eval=F, message=FALSE}
trc <- read.tree(paste(sep="",savedatafolder,"Data/squam_shl_new_Consensus_9755.tre"))

# Get evolutionary distinctiveness (fair proportion)
spp.ED <- evol.distinct(trc, type = c("fair.proportion"), ### Species' evolutionary distinctiveness
                               scale = FALSE, use.branch.lengths = TRUE)

# Get evolutionary distinctiveness (equal splits)
spp.ED_eq <- evol.distinct(trc, type = c("equal.splits"), ### Species' evolutionary distinctiveness
                               scale = FALSE, use.branch.lengths = TRUE)

Ncell <- nrow(occr)

cophe_tre <- cophenetic(tre)

bt<-branching.times(tre)

com_metrics <-
  pblapply(1:nrow(occr), FUN = function(x){
  
  tipnames <- which(occr[x,]==1) # species that occur in the cell i
  tipnames <- names(occr)[tipnames]

  trx <- drop.tip(tre, tre$tip.label[-which(tre$tip.label%in%tipnames)]) ### construct tree of all species within a cell

  subED<-subset(spp.ED, Species%in%tipnames)
  
  subED_eq<-subset(spp.ED_eq, Species%in%tipnames)
  
  SPD <- length(trx$tip.label) ### cell richness
  
  # Diversification rates as in Jetz et al.2012 (fair proportion)
  DIV <- mean(1/subED[,2]) 
  
  # Diversification rates as in Jetz et al.2012 (equal splits)
  DIV_eq <- mean(1/subED_eq[,2]) 
	
  # Diversification rates as in  Magallon & Sanderson 2001
  # Here using crown-group age
  DIV2 <- log(SPD)/(max(branching.times(trx))) 
  
  # Diversification rates as in  Magallon & Sanderson 2001
  # here using stem-group age
  foocrown<-findMRCA(tre, tipnames)
  stemnodemulti <- tre$edge[,1][tre$edge[,2]==foocrown]
  if(length(stemnodemulti)==0){
    stemage <- max(branching.times(trx))
  }
  else{
    stemage <- bt[names(bt)==stemnodemulti]
  }
	DIV3 <- log(SPD)/stemage
  
	### AGE
	# Crown-group age
	AGE <- max(branching.times(trx))
	# Stem-group age
	AGE2 <- stemage
	# max clade age
	clades_ <- getCladesofSize(trx, clade.size=2)
	AGE3 <- max(unlist(lapply(clades_, function(x) max(branching.times(x)))))
	# mean clade age
	clades_ <- getCladesofSize(trx, clade.size=2)
	AGE4 <- mean(unlist(lapply(clades_, function(x) max(branching.times(x)))))
	
  ### clade rich
  clades <- length(clades_) 
  
  # calc MPD
  # this is just for testing whether this MPD calculated by hand is equal to the one calculated using the picante library
  # for test use plo(MPD_, MPD)
  MPD_ <- mpd_b(cophe_tre, tipnames)
  
  return(data.frame(SPD,DIV,DIV_eq,DIV2,DIV3,
                    AGE,AGE2,AGE3,AGE4,
                    clades,MPD_))
})

com_metrics <- bind_rows(com_metrics)

```

# COMMUNITY METRICS SES
```{r community_metrics, eval=F, message=FALSE}

ses_MPD_realm<-ses_MPD_biome<-ses_MPD_biorealm<-rep(NA,Ncell)

for (i in 1:Ncell){
  cat("\r",i,"of", Ncell)
  
  # SES MPD Realm
  # we will be using the "richness" null model algorithm
  # the "independentswap" is more complicated
  if(is.na(Realm[i])){
  ses_MPD_realm[i] <- NA
  } 
  else {
    Realm.cells <- as.numeric(which(Realm==Realm[i])) # cells in the realm
    occ.temp <- occr[Realm.cells,] # realm subset
    sps.in.realm <- which(colSums(occ.temp)>0) # species that occur in the realm
    sps.in.realm <- names(occr)[sps.in.realm]
    
    # randomly select from the total pool of species in the realm [sps.in.realm] the same number of species observed in that cell.
    null.vals <- replicate(100, 
                           mpd_b(cophe_tre, 
                                 sample(sps.in.realm,com_metrics$SPD[i])))
    ses_MPD_realm[i] <- (com_metrics$MPD_[i] - mean(null.vals))/sd(null.vals)
  }
  
  # SES MPD Biorealm
  # we will be using the "richness" null model algorithm
  # the "independentswap" is more complicated
  if(is.na(Biorealm[i])){
  ses_MPD_biorealm[i] <- NA
  } 
  else {
    Biorealm.cells <- as.numeric(which(Biorealm==Biorealm[i])) # cells in the Biorealm
    occ.temp <- occr[Biorealm.cells,] # Biorealm subset
    sps.in.Biorealm <- which(colSums(occ.temp)>0) # species that occur in the Biorealm
    sps.in.Biorealm <- names(occr)[sps.in.Biorealm]
    
    # randomly select from the total pool of species in the Biorealm [sps.in.Biorealm] the same number of species observed in that cell.
    null.vals <- replicate(100, 
                           mpd_b(cophe_tre,
                                 sample(sps.in.Biorealm,com_metrics$SPD[i])))
    ses_MPD_biorealm[i] <- (com_metrics$MPD_[i] - mean(null.vals))/sd(null.vals)
  }
  
  # SES MPD biome
  # we will be using the "richness" null model algorithm
  # the "independentswap" is more complicated
  if(is.na(biome[i])){
  ses_MPD_biome[i] <- NA
  } 
  else {
    biome.cells <- as.numeric(which(biome==biome[i])) # cells in the biome
    occ.temp <- occr[biome.cells,] # biome subset
    sps.in.biome <- which(colSums(occ.temp)>0) # species that occur in the biome
    sps.in.biome <- names(occr)[sps.in.biome]
    
    # randomly select from the total pool of species in the biome [sps.in.biome] the same number of species observed in that cell.
    null.vals <- replicate(100, 
                           mpd_b(cophe_tre, 
                                 sample(sps.in.biome, com_metrics$SPD[i])))
    ses_MPD_biome[i] <- (com_metrics$MPD_[i] - mean(null.vals))/sd(null.vals)
  }
}

# we calculated MPDses using to the global species pool in order to get a metric that represents the community age relative to all species, not only those occurring at the regional species pool.
COM1 <- ses.mpd (occr, cophenetic(tre), null.model = "richness", runs=99, iterations=100)
MPD  <- COM1[,"mpd.obs"]
ses.MPD <- COM1[,"mpd.obs.z"]

COM2 <- ses.mpd (occr, cophenetic(tre), null.model = "independentswap", runs=99, iterations=100)
MPD2  <- COM2[,"mpd.obs"]
ses.MPD2 <- COM2[,"mpd.obs.z"]


com_metrics_ses <- data.frame(ses.MPD,ses.MPD2,
                              ses_MPD_realm,ses_MPD_biome,ses_MPD_biorealm)


gc()


```

# Functional diversity
```{r}
# Make a file with Nominal Variables
tabN <- data.frame(apply(trait[,c(3:8)], 2, as.factor))
# Now with the quantitative Variables
tabQ <- data.frame(trait[,c(1,2,9:11)])
rownames(tabQ) <- rownames(trait)
rownames(tabN) <- rownames(trait)

ktab1  <- ktab.list.df(list(tabN,tabQ))
TraitDis <- dist.ktab(ktab1 , c("N", "Q"), c("scaledBYrange"))

#FD <-dbFD(TraitDis, occr)

FD <- fdisp(TraitDis, as.matrix(occr))

str(FD)

## Calculate ses.FD
# ncores <- detectCores()
# cl <- makeCluster(ncores)
# registerDoParallel(cl)
# 
# obs.null.output <- foreach(i=1:100, .combine='cbind', .packages=c('FD','picante')) %dopar% { 
#   fdisp(TraitDis, randomizeMatrix(occr, null.model = "independentswap"))$FDis
# }
# # stop the cluster
# stopCluster(cl)
# 
# obs.null.output <- data.frame(obs.null.output)
# 
# ses.FD <- NA
# for(i in 1:Ncell){ cat("\r", i, 'from', dim(occr)[1])
#   ses.FD[i] <- (FD$FDis[i] - mean(as.numeric(obs.null.output[i,])))/sd(as.numeric(obs.null.output[i,]))
# }

###############
## Calculate ses.FD 
fdis_prepared <- fdis_prep(TraitDis, as.matrix(occr))

# Using Realms as species pools

## Calculate ses.FD Realm
ncores <- detectCores()
cl <- makeCluster(ncores)
registerDoParallel(cl)

FD.Realm.random <- foreach(i=1:Ncell, .combine='cbind', .packages=c('FD')) %dopar% { 
  # SES FDis
  # we will be using the "richness" null model algorithm
  # the "independentswap" is more complicated
  if(is.na(Realm[i])){
  c(rep(NA,100))
  } 
  else {
    Realm.cells <- as.numeric(which(Realm==Realm[i])) # cells in the realm
    occ.temp <- occr[Realm.cells,] # realm subset
    sps.in.realm <- which(colSums(occ.temp)>0) # species that occur in the realm
    sps.in.realm <- names(occr)[sps.in.realm]
    
    # randomly select from the total pool of species in the realm [sps.in.realm] the same number of species observed in that cell.
    null.vals <- replicate(100, 
                           tryCatch(
                             fdis_calc(fdis_prepared,
                                       sample(sps.in.realm,com_metrics$SPD[i]))$FDis,
                             error=function(err) NA))
    null.vals
  }
}
# stop the cluster
stopCluster(cl)

ses.FD.realm <- NA
for(i in 1:Ncell){ cat("\r", i, 'from', dim(occr)[1])
  ses.FD.realm[i] <- (FD$FDis[i] -
                        mean(as.numeric(na.omit(FD.Realm.random[,i])))/
    sd(as.numeric(na.omit(FD.Realm.random[,i]))))
}

# Using Biorealms as species pools

## Calculate ses.FD Biorealm
ncores <- detectCores()
cl <- makeCluster(ncores)
registerDoParallel(cl)

FD.Biorealm.random <- foreach(i=1:Ncell, .combine='cbind', .packages=c('FD')) %dopar% { 
  # SES FDis
  # we will be using the "richness" null model algorithm
  # the "independentswap" is more complicated
  if(is.na(Biorealm[i])){
  c(rep(NA,100))
  } 
  else {
    Biorealm.cells <- as.numeric(which(Biorealm==Biorealm[i])) # cells in the Biorealm
    occ.temp <- occr[Biorealm.cells,] # Biorealm subset
    sps.in.Biorealm <- which(colSums(occ.temp)>0) # species that occur in the Biorealm
    sps.in.Biorealm <- names(occr)[sps.in.Biorealm]
    
    # randomly select from the total pool of species in the Biorealm [sps.in.Biorealm] the same number of species observed in that cell.
    null.vals <- replicate(100, 
                           tryCatch(
                             fdis_calc(fdis_prepared,
                                       sample(sps.in.Biorealm,com_metrics$SPD[i]))$FDis,
                             error=function(err) NA))
    null.vals
  }
}
# stop the cluster
stopCluster(cl)

ses.FD.biorealm <- NA
for(i in 1:Ncell){ cat("\r", i, 'from', dim(occr)[1])
  ses.FD.biorealm[i] <- (FD$FDis[i] -
                        mean(as.numeric(na.omit(FD.Biorealm.random[,i])))/
    sd(as.numeric(na.omit(FD.Biorealm.random[,i]))))
}

# Using biome as species pools

## Calculate ses.FD biome
ncores <- detectCores()
cl <- makeCluster(ncores)
registerDoParallel(cl)

FD.biome.random <- foreach(i=1:Ncell, .combine='cbind', .packages=c('FD')) %dopar% { 
  # SES FDis
  # we will be using the "richness" null model algorithm
  # the "independentswap" is more complicated
  if(is.na(biome[i])){
  c(rep(NA,100))
  } 
  else {
    biome.cells <- as.numeric(which(biome==biome[i])) # cells in the biome
    occ.temp <- occr[biome.cells,] # biome subset
    sps.in.biome <- which(colSums(occ.temp)>0) # species that occur in the biome
    sps.in.biome <- names(occr)[sps.in.biome]
    
    # randomly select from the total pool of species in the biome [sps.in.biome] the same number of species observed in that cell.
    null.vals <- replicate(100, 
                           tryCatch(
                             fdis_calc(fdis_prepared,
                                       sample(sps.in.biome,com_metrics$SPD[i]))$FDis,
                             error=function(err) NA))
    null.vals
  }
}
# stop the cluster
stopCluster(cl)

ses.FD.biome <- NA
for(i in 1:Ncell){ cat("\r", i, 'from', dim(occr)[1])
  ses.FD.biome[i] <- (FD$FDis[i] -
                        mean(as.numeric(na.omit(FD.biome.random[,i])))/
    sd(as.numeric(na.omit(FD.biome.random[,i]))))
}


FD_metrics <- cbind(FDis=FD$FDis, ses.FD.realm, ses.FD.biorealm, ses.FD.biome)

```


# METADATA
```{r}
metadata <- cbind(data.frame(XY),
                  com_metrics,
                  com_metrics_ses,
                  FD_metrics,
                  env.vars)
```


# WRITE RESULTS
```{r}
write.csv(metadata, paste(savedatafolder,"Results/reptiles",'_metadata.csv',sep=""))
```


# SAVE
```{r echo=F, eval=FALSE}
save.image(paste(savedatafolder,"Code/reptiles",'_data.RData',sep=""))
```